Generating Images from Captions with Attention

### 2015-10-09 ram_workshop/ram-paper.tex
update

\vspace*{-1.0cm}
\title{Generating Images From Captions}

@@

\vspace{-0.1in}
\subsection{Image Modelling: the Conditional alignDRAW Network}
\comm{
The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step. While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables. 
}
To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}. The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step. Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}. Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
\begin{align}
\vspace{-0.1in}
\label{eq:x_hat}
\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
canv_t = \canv_{t-1} + \writeop(\hdec_t0, 
\label{eq:write}
%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}
where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
%\begin{align}
%\vspace{-0.1in}
$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
%\end{align}
The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
\comm{
\begin{align}
e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
\alpha_{j} &= \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}.
\end{align}
}
\comm{
Here $\hlang_{0}$ is initialized to the learned bias.
Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
}

@@

At the test time, images are generated by ancestral sampling the latent variables from the prior $\prior(\Lat_{1:T})$. 
%An additional post-processing step is performed to sharpen the often blurry images generated by the alignDRAW model. 
The generator of an adversarial network, trained independently on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW, which are often blurry. The adversarial network is trained as in \citep{denton_lapgan} but is conditioned on the skip-thought vector representation \cite{skip-thought} of the input caption.
Instead of sampling from its prior, we fix the input to the adversarial generator to be the mean of the original uniform distribution. This post-processing step is a deterministic mapping which enables us to calculate the lower bound on the new log-likelihood defined at the output of the adversarial generator. Interestingly, we found that the deterministic process generates samples with much less noise than if we had sampled from the uniform distribution.

>>>>

\vspace*{-1.0cm}
\title{Generating Images from Captions with Attention}

@@

\vspace{-0.1in}
\subsection{Image Modelling: the Conditional alignDRAW Network}
\comm{
The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step. While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables. 
}
To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}. The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step. Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}. Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
\begin{align}
\vspace{-0.1in}
\label{eq:x_hat}
\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
\canv_t = \canv_{t-1} + \writeop(\hdec_t), 
\label{eq:write}
%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}
where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}$:
%\begin{align}
%\vspace{-0.1in}
$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
%\end{align}
The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
\comm{
\begin{align}
e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
\alpha_{j} &= \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}.
\end{align}
}
\comm{
Here $\hlang_{0}$ is initialized to the learned bias.
Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
}

@@

At the test time, images are generated by ancestral sampling the latent variables from the prior $\prior(\Lat_{1:T})$. 
%An additional post-processing step is performed to sharpen the often blurry images generated by the alignDRAW model. 
The generator of an adversarial network, trained independently as in \cite{denton_lapgan} on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW, which are often blurry. 
%The adversarial network is trained as in \citep{denton_lapgan} but is conditioned on the skip-thought vector representation \cite{skip-thought} of the input caption.
Instead of sampling from its prior, we fix the input to the adversarial generator to be the mean of the original uniform distribution. This post-processing step is a deterministic mapping which enables us to calculate the lower bound on the new log-likelihood defined at the output of the adversarial generator. Interestingly, we found that the deterministic process generates samples with much less noise than if we had sampled from the uniform distribution.
